Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/evaluation.tex
fixed

New-Order presents two challenges: ID assignment and stock
maintainence. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with substantially more stock
than in a non-HAT-compliant implementation.

Delivery is challenging due to non-monotonicity. Each Delivery deletes
a pending order from the NewOrder Table and should be idempotent in
order to avoid billing a customer twice; this implies a need to
prevent Lost Update. We can avoid this issue by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that she is the only carrier who has
done so---but cannot provide a correct execution with HATs
alone. However, according to distributed transaction
architects~\cite{entitygroup}, these compensatory actions are
relatively common in real-world business processes.

@@

\begin{figure}[t!]
\begin{center}
\hspace{2em}\includegraphics[width=.8\columnwidth]{figs/strategylegend.pdf}
\end{center}\vspace{-2em}
\begin{center}\textbf{Within \texttt{us-east}}\end{center}\vspace{-1.5em}
\includegraphics[width=\columnwidth]{figs/lan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=\columnwidth]{figs/lan-threads-thru.pdf}
\begin{center}\textbf{Between \texttt{us-east} and \texttt{us-west-2}}\end{center}\vspace{-1.5em}
\includegraphics[width=\columnwidth]{figs/wan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=\columnwidth]{figs/wan-threads-thru.pdf}
\caption{YCSB performance for two clusters of three servers each
  deployed within a single datacenter and cross-datacenters.}
\label{fig:wan-exp}
\end{figure}

>>>>

New-Order presents two challenges: ID assignment and stock
maintainence. First, each New-Order transaction requires a unique ID
number for the order. We can create a unique number by, say,
concatenating the client ID and a timestamp. However, the TPC-C
specification requires order numbers to be \textit{sequentially}
assigned within a district, which requires preventing Lost
Update. Accordingly, HATs cannot provide compliant TPC-C execution but
can maintain uniqueness constraints. Second, the New-Order transaction
decrements inventory counts: what if the count becomes negative?
Fortunately, TPC-C New-Order restocks each item's inventory count
(increments by 91) if it would become negative as the result of
placing an order. This means that, even in the presence of concurrent
New-Orders, an item's stock will never fall below zero. This is TPC-C
compliant, but a HAT system might end up with much more stock
than in a non-HAT-compliant implementation.

Delivery is challenging due to non-monotonicity. Each Delivery deletes
a pending order from the NewOrder Table and should be idempotent in
order to avoid billing a customer twice; this implies a need to
prevent Lost Update. This issue  can be avoided  by moving the
non-monotonicity to the real world---the carrier that picks up the
package for an order can ensure that she is the only carrier who has
done so---but cannot provide a correct execution with HATs
alone. However, according to distributed transaction
architects~\cite{entitygroup}, these compensatory actions are
relatively common in real-world business processes.

@@

\begin{figure}[t!]
\begin{center}
\hspace{2em}\includegraphics[width=.8\columnwidth]{figs/strategylegend.pdf}
\end{center}\vspace{-2em}
\begin{center}\textbf{Within \texttt{us-east}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/lan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=0.90\columnwidth]{figs/lan-threads-thru.pdf}
\begin{center}\textbf{Between \texttt{us-east} and \texttt{us-west-2}}\end{center}\vspace{-1.5em}
\includegraphics[width=0.90\columnwidth]{figs/wan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=0.90\columnwidth]{figs/wan-threads-thru.pdf}
\caption{YCSB performance for two clusters of three servers each
  deployed within a single datacenter and cross-datacenters.}
\label{fig:wan-exp}
\end{figure}

### 2013-04-01 writeup/hat-analysis.tex
fixed

``Repeatable Read'' isolation is a confusing property. Some IBM
products use the term for fully serializable
isolation~\cite{hat-hotos}. Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all interpret
Repeatable Read as providing serializability for all operations except
predicate-based reads (and so Repeatable Read is identical to
serializability in a key-value store without predicate-based
access). In this model, the \textit{Phantom Problem}, whereby two
successive predicate-based reads return different
data~\cite{gray-isolation}, is the only kind of non-serializable
behavior allowed. When referring to ``Repeatable Read,'' we will
follow these definitions, which, as we will soon show, are for a
property that can be offered in a highly available way.

@@

There are several session guarantees that we can make with high
availability:

@@

\noindent{\textbf{Consistency}} A HAT system can make limited
consistency guarantees. It can often execute commutative and logically
monotonic~\cite{calm} operations without the risk of invalidating
(also monotonic) application-level integrity constraints. Our goal in
this paper is not to sketch the entire space of consistency models
that are achievable (see Section~\ref{sec:futurework}). We
specifically evaluate TPC-C transaction semantics under HAT
consistency guarantees in Section~\ref{sec:evaluation}.

@@

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:eval}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availabilty~\cite{gilbert-cap}.

@@

Composition cost also varies by combination. For instance, Charron-Bost
has proven that, to capture causality between $N$ communicating
processes, standard vector-based approaches face an upper bound of
$O(N)$ storage per write~\cite{charron-bost}. This means that, with
$100K$ clients, each write might be accompanied by $100K$ timestamps
per vector. This is difficult to scale. By compromising on
availability (e.g., treating a datacenter as a linearizable cluster),
this overhead can be reduced~\cite{eiger}, but it is much
cheaper to provide, say, read your writes, than full causal
consistency.

>>>>

``Repeatable Read'' isolation is a confusing property. Some IBM
products use the term for fully serializable
isolation~\cite{hat-hotos}. Gray~\cite{gray-isolation}, Berenson et
al.~\cite{ansicritique}, and Adya's PL-2.99~\cite{adya} all interpret
Repeatable Read as providing serializability for all operations except
for predicate-based reads (and so Repeatable Read is identical to
serializability in a key-value store without predicate-based
access). In this model, the \textit{Phantom Problem}, whereby two
successive predicate-based reads return different
data~\cite{gray-isolation}, is the only kind of non-serializable
behavior allowed. When referring to ``Repeatable Read,'' we will
follow these definitions, which, as we will soon show, are for a
property that can be offered in a highly available way.

@@

Several session guarantees can be made with high availability:

@@

\vspace{0.5em}
\noindent{\textbf{Consistency}} A HAT system can make limited
consistency guarantees. It can often execute commutative and logically
monotonic~\cite{calm} operations without the risk of invalidating
 application-level integrity constraints. Our goal in
this paper is not to sketch the entire space of consistency models
that are achievable (see Section~\ref{sec:futurework}). We
specifically evaluate TPC-C transaction semantics under HAT
consistency guarantees in Section~\ref{sec:evaluation}.

@@

It is impossible to prevent Lost Update in a highly available
environment. Consider two clients who submit the following $T_1$ and
$T_2$ on opposite sides of a network partition:
\begin{align*}
\small
T_1 &: r_x(100)~w_x(100+20=120)
\\T_2 &: r_x(100)~w_x(100+30=130)
\end{align*}
Regardless of whether $x=120$ or $x=130$ is chosen by a replica, the
database state could not have arisen serial execution of $T_1$ and
$T_2$.\footnote{In this example, we assume that, as is standard in
  modern databases, replicas accept values as they are written (i.e.,
  register semantics). This particular example could be made
  serializable via the use of commutative updates
  (Section~\ref{sec:evaluation}) but, as we show in the extended version of
  the paper~\cite{hat-tr}, the problem persists in the general case.}
To prevent this from happening, either $T_1$ or $T_2$ should not have
committed. Each client's respective server might try to detect that
another write occurred, but this requires knowing the version of the
latest write to $x$. In our example, this reduces to a requirement for
linearizability, which is, via Gilbert and Lynch's proof of the CAP
Theorem, provably unachievable with high
availabilty~\cite{gilbert-cap}.

@@

Composition cost also varies by combination. For instance, Charron-Bost
 proved that, to capture causality between $N$ communicating
processes, standard vector-based approaches face an upper bound of
$O(N)$ storage per write~\cite{charron-bost}. This means that, with
$100K$ clients, each write might be accompanied by $100K$ timestamps
per vector. This is difficult to scale. By compromising on
availability (e.g., treating a datacenter as a linearizable cluster),
this overhead can be reduced~\cite{eiger}, but it is much
cheaper to provide, say, read your writes, than full causal
consistency.
